Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2009-12-06 ltx/BrainIDE.tex
added MATLAB code of model

In order to implement standard estimation techniques we propose a decomposition of the The field using a set of Gaussian basis functions. Decomposition allows a continuous field to be represented by a finite dimensional state vector. This allows for the application of standard nonlinear, non-Gaussian state estimation methods such as sequential Monte-Carlo techniques like particle filtering. The field decomposition is described by
\begin{equation}\label{DefFieldDecomp}
{v_t}\left( {\bf{r}} \right) = {\phi ^T}\left( {\bf{r}} \right){x_t}.
\end{equation}
where $\phi(r)$ is the Gaussian basis functions that are scaled by the the state vector. The connectivity kernel can all be decomposed into a set of basis functions
\begin{equation}\label{DefKernelDecomp}
k( \mathbf{r} , \mathbf{r'} ) = \theta^T(\mathbf{r})\psi(\mathbf{r} , \mathbf{r'} )
\end{equation}
Will will assume we know the parametric form of the basis functions, where the parameter $\theta$ is unknown. Each connectivity basis function can be considered a layer in the WC model, representing short range excitation, surround inhibition and long range excitation. Making substitutions we get
\begin{equation}\label{DecompModel}
{\phi ^T}\left( \mathbf{r} \right){x_{t+1}} = \theta^T(\mathbf{r})\int_\Omega  {\psi \left( {{\bf{r}} , {\mathbf{r'}}} \right)f\left( {{\phi ^T}\left( {{\mathbf{r'}}} \right){x_t}} \right)d{\mathbf{r'}}}  - \alpha\phi^T(r)x_t + {\varepsilon _t}\left( {\mathbf{r}} \right).
\end{equation}
To isolate the state we cross multiply by $\phi(r)$ and integrate over $\bf{r}$
\begin{equation}\label{Poo}
\int_\Omega  {\phi \left( {\bf{r}} \right){\phi ^T}\left( {\bf{r}} \right)d{\bf{r}}{x_{t+1}}}  = \int_\Omega  {\phi ( \mathbf{r} )\theta^T(\mathbf{r})\int_\Omega  {\psi \left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}} d{\bf{r}}}  - \alpha\int_\Omega \phi(r)\phi^T(r)d\bf{r} x_t  + \int_\Omega  {\phi \left( {\bf{r}} \right){\varepsilon _t}\left( {\bf{r}} \right)d{\bf{r}}}.
\end{equation}
Now we define
\begin{equation}\label{DefGamma}
\Gamma  = \int_\Omega  {\phi \left( {\bf{r}} \right){\phi ^T}\left( {\bf{r}} \right)d{\bf{r}}}.
\end{equation}
Substituting equation~\ref{DefGamma} into \ref{Poo} and rearranging gives
\begin{equation}\label{SS Model}
{x_{t+1}} = {\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right)\theta^T(\mathbf{r})\int_\Omega  {\psi \left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}} d{\bf{r}}}-\alpha x_t + {\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right){\varepsilon _t}\left( {\bf{r}} \right)d{\bf{r}}}.
\end{equation}
Under the assumption of spatial homogeneity we can write
\begin{equation}\label{Homogeneous SS Model}
{x_{t+1}} = \theta^T{\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right)\int_\Omega  {\psi \left( {{\bf{r}} - {\bf{r'}}} \right)f\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}} d{\bf{r}}}-\alpha x_t + {\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right){\varepsilon _t}\left( {\bf{r}} \right)d{\bf{r}}}.
\end{equation}
This provides the final state-space model
\begin{equation}\label{AbbrevSSModel}
x_{t + 1} = \theta^T q(x) -\alpha x + \Gamma ^{ - 1}\int_\Omega  {\phi ( \mathbf{r} ){\varepsilon _t}( \mathbf{r} )d\mathbf{r}}
\end{equation}
with observation equation
\begin{equation}\label{ObservationEquation}
    y_t( k ) = \int_\Omega  m( k,\mathbf{r} )\phi^T( \mathbf{r} ) d\mathbf{r}x_t  + \varepsilon ( k ),
\end{equation}

>>>>

In order to implement standard estimation techniques we propose a decomposition of the The field using a set of Gaussian basis functions. Decomposition allows a continuous field to be represented by a finite dimensional state vector. This allows for the application of standard nonlinear, non-Gaussian state estimation methods such as sequential Monte-Carlo techniques like particle filtering. The field decomposition is described by
\begin{equation}\label{DefFieldDecomp}
{v_t}\left( {\bf{r}} \right) = {\phi ^T}\left( {\bf{r}} \right){x_t}.
\end{equation}
where $\phi(r)$ is the Gaussian basis functions that are scaled by the the state vector. The connectivity kernel can all be decomposed into a set of basis functions
\begin{equation}\label{DefKernelDecomp}
k( \mathbf{r} , \mathbf{r'} ) = \theta^T\psi(\mathbf{r} , \mathbf{r'} )
\end{equation}
Will will assume we know the parametric form of the basis functions, where the parameter $\theta$ is unknown. Each connectivity basis function can be considered a layer in the WC model, representing short range excitation, surround inhibition and long range excitation. Making substitutions we get
\begin{equation}\label{DecompModel}
{\phi ^T}\left( \mathbf{r} \right){x_{t+1}} = \theta^T\int_\Omega  {\psi \left( {{\bf{r}} , {\mathbf{r'}}} \right)f\left( {{\phi ^T}\left( {{\mathbf{r'}}} \right){x_t}} \right)d{\mathbf{r'}}}  - \alpha\phi^T(\mathbf{r})x_t + {\varepsilon _t}\left( {\mathbf{r}} \right).
\end{equation}
To isolate the state we cross multiply by $\phi(\mathbf{r})$ and integrate over $\mathbf{r}$
\begin{equation}\label{Poo}
\int_\Omega  {\phi \left( {\bf{r}} \right){\phi ^T}\left( {\bf{r}} \right)d{\bf{r}}{x_{t+1}}}  = \int_\Omega  {\phi ( \mathbf{r} )\theta^T\int_\Omega  {\psi \left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}} d{\bf{r}}}  - \alpha\int_\Omega \phi(\mathbf{r})\phi^T(\mathbf{r})d\mathbf{r} x_t  + \int_\Omega  {\phi \left( {\mathbf{r}} \right){\varepsilon _t}\left( {\mathbf{r}} \right)d{\mathbf{r}}}.
\end{equation}
Now we define
\begin{equation}\label{DefGamma}
\Gamma  = \int_\Omega  {\phi \left( {\bf{r}} \right){\phi ^T}\left( {\bf{r}} \right)d{\bf{r}}}.
\end{equation}
Substituting equation~\ref{DefGamma} into \ref{Poo} and rearranging gives
\begin{equation}\label{SS Model}
{x_{t+1}} = {\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right)\theta^T\int_\Omega  {\psi \left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}} d{\bf{r}}}-\alpha x_t + {\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right){\varepsilon _t}\left( {\bf{r}} \right)d{\bf{r}}}.
\end{equation}
We can rearrange to get
\begin{equation}\label{Homogeneous SS Model}
{x_{t+1}} = {\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right)\int_\Omega  {f\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)\psi^T \left( {{\bf{r}} - {\bf{r'}}} \right)d{\bf{r'}}} d{\bf{r}}}\theta - \alpha x_t + {\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right){\varepsilon _t}\left( {\bf{r}} \right)d{\bf{r}}}.
\end{equation}
This provides the final state-space model
\begin{equation}\label{AbbrevSSModel}
x_{t + 1} = \theta^T q(x) -\alpha x + \Gamma ^{ - 1}\int_\Omega  {\phi ( \mathbf{r} ){\varepsilon _t}( \mathbf{r} )d\mathbf{r}}
\end{equation}
with observation equation
\begin{equation}\label{ObservationEquation}
    y_t( k ) = \int_\Omega  m( k,\mathbf{r} )\phi^T( \mathbf{r} ) d\mathbf{r}x_t  + \varepsilon ( k ),
\end{equation}
